Formalization, Mechanization and Automation of G\"{o}del's Proof of God's Existence\thanks{This work has been supported by the German Research Foundation under grant BE2501/9-1. Bruno ????}

### 2013-08-19 Abstract/GodProof-Abstract.tex

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  G\"{o}del's proof \cite{Goedel1970} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
%
G\"{o}del defines God as a being who possesses all \emph{positive} properties.
He does not extensively discuss what positive properties are, 
but instead he states a few reasonable but debatable axioms that they should satisfy.
Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof (e.g. \cite{Scott,Sobel,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). Our formalization employs the following axioms (A*) and definitions (D*):\footnote{
(A1), (A2), (A5), (D1), (D3) are logically equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and 3 in G\"odel's manuscript \cite{Goedel1970}. (A3) was introduced by Scott \cite{Scott} and could be derived from G\"odel's axiom 1 and (D1) in a logic with infinitary conjunction. (A4) is a weaker form of G\"odel's axiom 3. (D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
}
\marginpar{\scriptsize Chris: The reference to Scott's version is kind of confusing, we need to verify this source.}
\marginpar{\scriptsize Bruno: I think the numbers (A1, A2,.., D1,..D3, L1,..,L3,T) should be in the left with a description environment, as they were before. I think it looked nicer like that. If you agree, I can change back.}
\allowdisplaybreaks[1] 
\begin{align}
& \text{Any property necessarily implied by a positive property is positive.} \notag \\
& \quad 
\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)] \tag{A1} \\
& \text{A property is positive if and only if its negation is not positive.} \notag \\
& \quad 
\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
\tag{A2} \\
& \text{The property of being God-like is positive.} \notag \\
& \quad 
P(G) 
\tag{A3} \\
& \text{Positive properties are necessarily positive.} \notag \\
& \quad 
\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
\tag{A4} \\
& \text{Necessary existence is a positive property.} \notag \\
& \quad 
P(E)
\tag{A5} \\
& \text{x is God-like if and only if x has every positive property.} \notag \\
& \quad 
G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
\tag{D1} \\ 
& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P.}\notag \\
& \quad 
\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all x. (\varphi(x) \imp \psi(x)))
\tag{D2} \\
& \parbox[c][2em][c]{0.8\textwidth}{x necessarily exists if and only if every essence of x is necessarily exemplified.} \notag \\
& \quad 
E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex x.\varphi(x)] 
\tag{D3} 
\end{align}

\noindent
From these axioms and definitions we then infer:
\begin{align}
& \text{Positive properties are possibly exemplified.} \notag \\
& \quad 
\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
\tag{L1} \\
& \text{Possibly God exists.} \notag \\
& \quad 
\pos \ex x. G(x)
\tag{L2} \\
& \parbox[c][2em][c]{0.8\textwidth}{If x is God-like, then the property of being God-like is an essence of x.} \notag \\
& \quad 
\all y.[G(y) \imp \ess{G}{y}]
\tag{L3} \\
& \text{Necessarily God exists.} \notag \\
& \quad 
\nec \ex x. G(x)
\tag{T}
\end{align}

>>>>

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  G\"{o}del's proof \cite{Goedel1970} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
%
G\"{o}del defines God as a being who possesses all \emph{positive} properties.
He does not extensively discuss what positive properties are, 
but instead he states a few reasonable but debatable axioms that they should satisfy.
Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof (e.g. \cite{Scott,Sobel,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). Our formalization employs the following axioms (A*) and definitions (D*):\footnote{
(A1), (A2), (A5), (D1), (D3) are logically equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and 3 in G\"odel's manuscript \cite{Goedel1970}. (A3) was introduced by Scott \cite{Scott} and could be derived from G\"odel's axiom 1 and (D1) in a logic with infinitary conjunction. (A4) is a weaker form of G\"odel's axiom 3. (D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
}
\marginpar{\scriptsize Chris: The reference to Scott's version is kind of confusing, we need to verify this source.}
\marginpar{\scriptsize Bruno: I think the numbers (A1, A2,.., D1,..D3, L1,..,L3,T) should be in the left with a description environment, as they were before. I think it looked nicer before. Can I change it back?}
\allowdisplaybreaks[1] 
\begin{align}
& \text{Any property necessarily implied by a positive property is positive.} \notag \\
& \quad 
\all \varphi. \all \psi.[(P(\varphi) \wedge \nec \all x.[\varphi(x) \imp \psi(x)]) \imp P(\psi)] \tag{A1} \\
& \text{A property is positive if and only if its negation is not positive.} \notag \\
& \quad 
\all \varphi. [P(\neg \varphi) \biimp \neg P(\varphi)]
\tag{A2} \\
& \text{The property of being God-like is positive.} \notag \\
& \quad 
P(G) 
\tag{A3} \\
& \text{Positive properties are necessarily positive.} \notag \\
& \quad 
\all \varphi.[P(\varphi) \to \Box \; P(\varphi)]
\tag{A4} \\
& \text{Necessary existence is a positive property.} \notag \\
& \quad 
P(E)
\tag{A5} \\
& \text{Being God-like is having every positive property.} \notag \\
& \quad 
G(x) \biimp \forall \varphi. [P(\varphi) \to \varphi(x)]
\tag{D1} \\ 
& \parbox[c][2em][c]{0.8\textwidth}{A property is an essence of an individual if and only if the individual has it and every property  that the individual has is necessarily implied by it.}\notag \\
& \quad 
\ess{\varphi}{x} \biimp \varphi(x) \wedge \all \psi. (\psi(x) \imp \nec \all y. (\varphi(y) \imp \psi(y)))
\tag{D2} \\
& \parbox[c][2em][c]{0.8\textwidth}{An individual necessarily exists if and only if all its essences are necessarily exemplified.} \notag \\
& \quad 
E(x) \biimp \all \varphi.[\ess{\varphi}{x} \imp \nec \ex x.\varphi(x)] 
\tag{D3} 
\end{align}

\noindent
From these axioms and definitions we then infer:
\begin{align}
& \text{Positive properties are possibly exemplified.} \notag \\
& \quad 
\all \varphi. [P(\varphi) \imp \pos \ex x.\varphi(x)]
\tag{L1} \\
& \text{Possibly, God exists.} \notag \\
& \quad 
\pos \ex x. G(x)
\tag{L2} \\
& \parbox[c][2em][c]{0.8\textwidth}{The property of being God-like is an essence of any God-like being.} \notag \\
& \quad 
\all y.[G(y) \imp \ess{G}{y}]
\tag{L3} \\
& \text{Necessarily, God exists.} \notag \\
& \quad 
\nec \ex x. G(x)
\tag{T}
\end{align}
\marginpar{\scriptsize Bruno: I think it might be possible to improve the natural language statements further. After our improvements converge here, we should update the statements in all formalizations (THF, Coq, Isabelle, ND).}
